---
title: "Diabetes-Eye Exam Analysis"
author: "Mridula (Mally) Shan - Summer 2024"
output: pdf_document
urlcolor: blue
---

# Introduction & Code Set-Up

```{r, message = FALSE, warning = FALSE}

library(tidyverse)
library(dplyr)
library(infer)
library(readr)
library(broom)
library(gglm)
library(MASS)
library(AICcmodavg)
library(finalfit)
library(knitr)
library(lme4)
library(tinytex)

# (For Reference) Code to Reformat the Data for Analysis
# Collected data from the following link: 
#https://www.cdc.gov/nchs/nhis/2022nhis.htm
# surveying 27,651 people throughout the United States 

NHIS_Diabetes_Eye <- read_csv("adult22.csv")

```

# Descriptive Statistics 

### 1) Number Respondents from Various Regions of US

```{r, message = FALSE, warning = FALSE}

geographic_summary <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(REGION = case_when(REGION == 1 ~ "Northeast",
                            REGION == 2 ~ "Midwest",
                            REGION == 3 ~ "South", 
                            REGION == 4 ~ "West")) %>%
  dplyr::count(REGION) %>%
  dplyr::mutate(perc_region = (n/sum(n)*100)) %>%
  arrange(desc(perc_region))

kable(geographic_summary)

geographic_summary %>%
  ggplot(aes(x = reorder(REGION, -n), y = n)) + 
  geom_bar(stat = "identity", fill = "darkcyan") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number Respondents from Various Regions of US ",
       x ="Regions of US", 
       y = "Count")

```

### 2) Number Respondents from Various County Types in US 

The 2013 NCHS Urban-Rural Classification scheme defines: 

- **Large central metro** as counties (or county equivalents) in metropolitan statistical areas (MSA) of 1 million or more population that 1) contain the entire population of the largest principal city of the MSA, or 2) are completely contained in the largest principal city of the MSA, or 3) contain at least 250,000 residents of any principal city of the MSA.  

- **Large fringe metro** are counties (or county equivalents) in MSAs of 1 million or more
population that do not qualify as large central. 

- **Medium and small metro** are counties (or county equivalents) in MSAs of 250,000 to 999,999 population or in MSAs of less than 250,000 population. 

- **Nonmetropolitan** are counties (or county equivalents) in micropolitan statistical areas and non-core counties.

```{r, message = FALSE, warning = FALSE}

urban_summary <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(URBRRL = case_when(URBRRL == 1 ~ "Large central metro",
                                   URBRRL == 2 ~ "Large fringe metro",
                                   URBRRL == 3 ~ "Medium and small metro", 
                                   URBRRL == 4 ~ "Nonmetropolitan")) %>%
  dplyr::count(URBRRL) %>%
  dplyr::mutate(perc_urban = (n/sum(n)*100)) %>%
  arrange(desc(perc_urban))

kable(urban_summary)

urban_summary %>%
  ggplot(aes(x = reorder(URBRRL, -n), y = n)) + 
  geom_bar(stat = "identity", fill = "darkcyan") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number Respondents from Various County Types in US ",
       x ="Regions of US", 
       y = "Count")

```

### 3) Distribution of Respondents' Sex

```{r, message = FALSE, warning = FALSE}

dist_sex <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(SEX_A = case_when(SEX_A == 1 ~ "Male",
                                  SEX_A == 2 ~ "Female",
                                  SEX_A == 7 ~ "Refused", 
                                  SEX_A == 8 ~ "Not Ascertained",
                                  SEX_A == 9 ~ "Don't Know")) %>%
  dplyr::count(SEX_A) %>%
  dplyr::mutate(perc_sex = (n/sum(n)*100)) %>%
  arrange(desc(perc_sex))

kable(dist_sex)

dist_sex %>%
  ggplot(aes(x = reorder(SEX_A, -n), y = n)) + 
  geom_bar(stat = "identity", fill = "darkcyan") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title= "Distribution of Respondents' Sex",
       x ="Respondents' Sex", 
       y = "Count")

```

### 4) Distribution of Respondents' Race 

```{r, message = FALSE, warning = FALSE}

dist_race <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(RACEALLP_A = case_when(HISPALLP_A == 1 ~ "Hispanic",
                                       HISPALLP_A == 2 ~ "White only",
                                       HISPALLP_A == 3 ~ "Black/African American only", 
                                       HISPALLP_A == 4 ~ "Asian only",
                                       HISPALLP_A == 5 ~ "AIAN only",
                                       HISPALLP_A == 6 ~ "AIAN and any other group",
                                       HISPALLP_A == 7 ~ "Other single and multiple races", 
                                       HISPALLP_A == 97 ~ "Refused",
                                       HISPALLP_A == 98 ~ "Not Ascertained",
                                       HISPALLP_A == 98 ~ "Don't know")) %>%
  dplyr::count(RACEALLP_A) %>%
  dplyr::mutate(perc_race = (n/sum(n)*100)) %>%
  arrange(desc(perc_race))

kable(dist_race)

dist_race %>%
  ggplot(aes(x = reorder(RACEALLP_A, -n), y = n)) + 
  geom_bar(stat = "identity", fill = "darkcyan") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Distribution of Respondents' Race ",
       x ="Respondents' Race", 
       y = "Count")

```

### 5) Number of Respondents that had an Eye Exam

```{r, message = FALSE, warning = FALSE}

num_eye_exam <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(EYEEX12M_A = case_when(EYEEX12M_A == 1 ~ "Yes",
                                       EYEEX12M_A == 2 ~ "No",
                                       EYEEX12M_A == 7 ~ "Refused", 
                                       EYEEX12M_A == 8 ~ "Not Ascertained",
                                       EYEEX12M_A == 9 ~ "Don't Know")) %>%
  dplyr::count(EYEEX12M_A) %>%
  dplyr::mutate(perc_eye_exam = (n/sum(n)*100)) %>%
  arrange(desc(perc_eye_exam))

kable(num_eye_exam)

num_eye_exam %>%
  ggplot(aes(x = reorder(EYEEX12M_A, -n), y = n)) + 
  geom_bar(stat = "identity", fill = "darkcyan") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of Respondents that had an Eye Exam",
       x ="Did you have an eye exam?", 
       y = "Count")

```

### 6) Number of Respondents that have Diabetes 

```{r, message = FALSE, warning = FALSE}

num_diabetes <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                    DIBEV_A == 2 ~ "No",
                                    DIBEV_A == 7 ~ "Refused", 
                                    DIBEV_A == 8 ~ "Not Ascertained",
                                    DIBEV_A == 9 ~ "Don't Know")) %>%
  dplyr::count(DIBEV_A) %>%
  dplyr::mutate(perc_diabetes = (n/sum(n)*100)) %>%
  arrange(desc(perc_diabetes))

kable(num_diabetes)

num_diabetes %>%
  ggplot(aes(x = reorder(DIBEV_A, -n), y = n)) + 
  geom_bar(stat = "identity", fill = "darkcyan") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of Respondents that have Diabetes",
       x ="Did you ever have diabetes?", 
       y = "Count")

```
\newpage 

# Exploratory Analysis 

### 1) Number of Respondents with Diabetes who got an Eye Exam

```{r, message = FALSE, warning = FALSE}

has_diabetes_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(EYEEX12M_A = case_when(EYEEX12M_A == 1 ~ "Yes",
                                       EYEEX12M_A == 2 ~ "No",
                                       EYEEX12M_A == 7 ~ "Refused", 
                                       EYEEX12M_A == 8 ~ "Not Ascertained",
                                       EYEEX12M_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                       DIBEV_A == 2 ~ "No",
                                       DIBEV_A == 7 ~ "Refused", 
                                       DIBEV_A == 8 ~ "Not Ascertained",
                                       DIBEV_A == 9 ~ "Don't Know")) %>%
  filter(EYEEX12M_A %in% c("Yes", "No"), DIBEV_A %in% c("Yes", "No")) %>%
  dplyr::group_by(DIBEV_A) %>%
  dplyr::count(EYEEX12M_A) %>%
  dplyr::mutate(perc_eye_exam = (n/sum(n)*100)) %>%
  rename(Eye_Exam = EYEEX12M_A,
         Has_Diabetes = DIBEV_A)
  

kable(has_diabetes_tbl)

NHIS_Diabetes_Eye %>%
  dplyr::mutate(EYEEX12M_A = case_when(EYEEX12M_A == 1 ~ "Yes",
                                       EYEEX12M_A == 2 ~ "No",
                                       EYEEX12M_A == 7 ~ "Refused", 
                                       EYEEX12M_A == 8 ~ "Not Ascertained",
                                       EYEEX12M_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                       DIBEV_A == 2 ~ "No",
                                       DIBEV_A == 7 ~ "Refused", 
                                       DIBEV_A == 8 ~ "Not Ascertained",
                                       DIBEV_A == 9 ~ "Don't Know")) %>%
  filter(EYEEX12M_A %in% c("Yes", "No"), DIBEV_A %in% c("Yes")) %>%
  ggplot() + 
  geom_bar(mapping = aes(x = EYEEX12M_A),
           fill = "darkcyan") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of Respondents with Diabetes who got an Eye Exam",
       x ="Eye Exam Response", 
       y = "Count")


```


### 2) Number of Respondents with Gestational Diabetes who got an Eye Exam

``` {r, message = FALSE, warning = FALSE}

#Gestational diabetes
gestational_type_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(EYEEX12M_A = case_when(EYEEX12M_A == 1 ~ "Yes",
                                       EYEEX12M_A == 2 ~ "No",
                                       EYEEX12M_A == 7 ~ "Refused", 
                                       EYEEX12M_A == 8 ~ "Not Ascertained",
                                       EYEEX12M_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(GESDIB_A = case_when(GESDIB_A == 1 ~ "Yes",
                                     GESDIB_A == 2 ~ "No",
                                     GESDIB_A == 7 ~ "Refused", 
                                     GESDIB_A == 8 ~ "Not Ascertained",
                                     GESDIB_A == 9 ~ "Don't Know")) %>%
  filter(EYEEX12M_A %in% c("Yes", "No"), GESDIB_A %in% c("Yes")) %>%
  dplyr::group_by(GESDIB_A) %>%
  dplyr::count(EYEEX12M_A) %>%
  dplyr::mutate(perc_eye_exam = (n/sum(n)*100)) %>%
  rename(Eye_Exam = EYEEX12M_A,
         Gestational_Diabetes = GESDIB_A)

kable(gestational_type_tbl)


NHIS_Diabetes_Eye %>%
  dplyr::mutate(EYEEX12M_A = case_when(EYEEX12M_A == 1 ~ "Yes",
                                       EYEEX12M_A == 2 ~ "No",
                                       EYEEX12M_A == 7 ~ "Refused", 
                                       EYEEX12M_A == 8 ~ "Not Ascertained",
                                       EYEEX12M_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(GESDIB_A = case_when(GESDIB_A == 1 ~ "Yes",
                                     GESDIB_A == 2 ~ "No",
                                     GESDIB_A == 7 ~ "Refused", 
                                     GESDIB_A == 8 ~ "Not Ascertained",
                                     GESDIB_A == 9 ~ "Don't Know")) %>%
  filter(EYEEX12M_A %in% c("Yes", "No"), GESDIB_A %in% c("Yes")) %>%
  ggplot(mapping = aes(x = EYEEX12M_A)) + 
  geom_bar(fill = "darkcyan") +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Number of Respondents with Gestational Diabetes who got an Eye Exam",
       x ="Eye Exam Response", 
       y = "Count")

```


\newpage 

# Latest Findings 

## 1) Looking at differences in responses based on type of diabetes 

### a) Proportion of respondents who got an Eye Exam based on Diabetes type

```{r, message = FALSE, warning = FALSE}

diabetes_type_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(EYEEX12M_A = case_when(EYEEX12M_A == 1 ~ "Yes",
                                       EYEEX12M_A == 2 ~ "No",
                                       EYEEX12M_A == 7 ~ "Refused", 
                                       EYEEX12M_A == 8 ~ "Not Ascertained",
                                       EYEEX12M_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(DIBTYPE_A = case_when(DIBTYPE_A == 1 ~ "Type 1",
                                      DIBTYPE_A == 2 ~ "Type 2",
                                      DIBTYPE_A == 3 ~ "Other type of diabetes",
                                      DIBTYPE_A == 7 ~ "Refused", 
                                      DIBTYPE_A == 8 ~ "Not Ascertained",
                                      DIBTYPE_A == 9 ~ "Don't Know")) %>%
  filter(EYEEX12M_A %in% c("Yes", "No"), 
         DIBTYPE_A %in% c("Type 1", "Type 2", "Other type of diabetes")) %>%
  dplyr::group_by(DIBTYPE_A) %>%
  dplyr::count(EYEEX12M_A) %>%
  dplyr::mutate(perc_eye_exam = (n/sum(n)*100)) %>%
  rename(Diabetes_Type = DIBTYPE_A,
         Eye_Exam = EYEEX12M_A)

kable(diabetes_type_tbl)

NHIS_Diabetes_Eye %>%
  dplyr::mutate(EYEEX12M_A = case_when(EYEEX12M_A == 1 ~ "Yes",
                                       EYEEX12M_A == 2 ~ "No",
                                       EYEEX12M_A == 7 ~ "Refused", 
                                       EYEEX12M_A == 8 ~ "Not Ascertained",
                                       EYEEX12M_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(DIBTYPE_A = case_when(DIBTYPE_A == 1 ~ "Type 1",
                                      DIBTYPE_A == 2 ~ "Type 2",
                                      DIBTYPE_A == 3 ~ "Other type of diabetes",
                                      DIBTYPE_A == 7 ~ "Refused", 
                                      DIBTYPE_A == 8 ~ "Not Ascertained",
                                      DIBTYPE_A == 9 ~ "Don't Know")) %>%
  filter(EYEEX12M_A %in% c("Yes", "No"), 
         DIBTYPE_A %in% c("Type 1", "Type 2", "Other type of diabetes")) %>%
  ggplot(mapping = aes(x = DIBTYPE_A, fill = EYEEX12M_A)) + 
  geom_bar(position = "fill") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  scale_fill_manual(values = c("darkseagreen3", 
                               "darkcyan")) +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Proportion of respondents who got an Eye Exam based on Diabetes Type",
       x ="Eye Exam Response", 
       y = "Count")


```

## 2) Looking at differences in responses based on respondents' sex

### a) Number of respondents that have diabetes across the sexes 

```{r, message = FALSE, warning = FALSE}

diabetes_sex_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(SEX_A = case_when(SEX_A == 1 ~ "Male",
                                       SEX_A == 2 ~ "Female",
                                       SEX_A == 7 ~ "Refused", 
                                       SEX_A == 8 ~ "Not Ascertained",
                                       SEX_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                      DIBEV_A == 2 ~ "No",
                                      DIBEV_A == 7 ~ "Refused", 
                                      DIBEV_A == 8 ~ "Not Ascertained",
                                      DIBEV_A == 9 ~ "Don't Know")) %>%
  filter(DIBEV_A %in% c("Yes", "No"), 
         SEX_A %in% c("Male", "Female")) %>%
  dplyr::group_by(DIBEV_A) %>%
  dplyr::count(SEX_A) %>%
  dplyr::mutate(perc_diabetes = (n/sum(n)*100)) %>%
  rename(Respondent_Sex = SEX_A,
         Respondent_Diabetes = DIBEV_A)

kable(diabetes_sex_tbl)

diabetes_sex_tbl %>%
  ggplot(aes(x = reorder(Respondent_Diabetes, -n), y = n, fill = Respondent_Sex)) + 
  geom_col(position = "fill") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of respondents that have diabetes across the sexes",
       x ="Does the respondent have diabetes?", 
       y = "Proportion")



diabetes_sex_tbl %>%
  ggplot(aes(x = reorder(Respondent_Diabetes, -n), y = n, fill = Respondent_Sex)) + 
  geom_col(position = "dodge2") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of respondents that have diabetes across the sexes",
       x ="Does the respondent have diabetes?", 
       y = "Count")


```

### b) Number of respondents with diabetes that took an eye exam across the sexes

```{r}

eye_exam_sex_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(SEX_A = case_when(SEX_A == 1 ~ "Male",
                                       SEX_A == 2 ~ "Female",
                                       SEX_A == 7 ~ "Refused", 
                                       SEX_A == 8 ~ "Not Ascertained",
                                       SEX_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                      DIBEV_A == 2 ~ "No",
                                      DIBEV_A == 7 ~ "Refused", 
                                      DIBEV_A == 8 ~ "Not Ascertained",
                                      DIBEV_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(EYEEX12M_A =case_when (EYEEX12M_A == 1 ~"Yes",
                                       EYEEX12M_A == 2 ~"No",
                                       EYEEX12M_A == 7 ~"Refused",
                                       EYEEX12M_A == 8 ~"Not Ascertained",
                                       EYEEX12M_A == 9~"Don't Know")) %>%
  filter(DIBEV_A %in% c("Yes"), 
         SEX_A %in% c("Male", "Female"),
         EYEEX12M_A%in% c("Yes", "No")) %>%
  dplyr::group_by(EYEEX12M_A) %>%
  dplyr::count(SEX_A) %>%
  dplyr::mutate(perc_eye_exam = (n/sum(n)*100)) %>%
  rename(Respondent_Sex = SEX_A,
         Eye_Exam = EYEEX12M_A)

knitr::kable(eye_exam_sex_tbl)

eye_exam_sex_tbl %>%
  ggplot(aes(x = reorder(Eye_Exam, -n), y = n, fill = Respondent_Sex)) + 
  geom_col(position = "fill") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Number of respondents with diabetes that took an eye exam across the sexes",
       x ="Did the respondent take an eye exam?", 
       y = "Proportion")

eye_exam_sex_tbl %>%
  ggplot(aes(x = reorder(Eye_Exam, -n), y = n, fill = Respondent_Sex)) + 
  geom_col(position = "dodge2") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Number of respondents with diabetes that took an eye exam across the sexes",
       x ="Did the respondent take an eye exam?", 
       y = "Count")

```


## 3) Looking at differences in responses based on respondents' race 

### a) Number of respondents that have diabetes across races

```{r, message = FALSE, warning = FALSE}

diabetes_race_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(HISPALLP_A = case_when(HISPALLP_A == 1 ~ "Hispanic",
                                       HISPALLP_A == 2 ~ "White only",
                                       HISPALLP_A == 3 ~ "Black/African American only", 
                                       HISPALLP_A == 4 ~ "Asian only",
                                       HISPALLP_A == 5 ~ "AIAN only",
                                       HISPALLP_A == 6 ~ "AIAN and any other group",
                                       HISPALLP_A == 7 ~ "Other single and multiple races",
                                       HISPALLP_A == 97 ~ "Refused",
                                       HISPALLP_A == 98 ~ "Not Ascertained",
                                       HISPALLP_A == 99 ~ "Don't Know")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                      DIBEV_A == 2 ~ "No",
                                      DIBEV_A == 7 ~ "Refused", 
                                      DIBEV_A == 8 ~ "Not Ascertained",
                                      DIBEV_A == 9 ~ "Don't Know")) %>%
  filter(DIBEV_A %in% c("Yes", "No"), 
         HISPALLP_A %in% c("Hispanic", "White only", "Black/African American only",
                      "Asian only", "AIAN only", "AIAN and any other group",
                      "Other single and multiple races")) %>%
  dplyr::group_by(DIBEV_A) %>%
  dplyr::count(HISPALLP_A) %>%
  dplyr::mutate(perc_diabetes = (n/sum(n)*100)) %>%
  rename(Respondent_Race = HISPALLP_A,
         Respondent_Diabetes = DIBEV_A)

kable(diabetes_race_tbl)

diabetes_race_tbl %>%
  ggplot(aes(x = reorder(Respondent_Diabetes, -n), y = n, fill = Respondent_Race)) + 
  geom_col(position = "fill") + 
  scale_fill_brewer(palette = "Paired") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of respondents that have diabetes across the races",
       x ="Does the respondent have diabetes?", 
       y = "Proportion")

diabetes_race_tbl %>%
  ggplot(aes(x = reorder(Respondent_Diabetes, +n), y = n, fill = Respondent_Race)) + 
  geom_col(position = "dodge2") + 
  scale_fill_brewer(palette = "Paired") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of respondents that have diabetes across the races",
       x ="Does the respondent have diabetes?", 
       y = "Count")

```

### b) Number of respondents with diabetes that took an eye exam across races

```{r}


eye_exam_race_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(HISPALLP_A = case_when(HISPALLP_A == 1 ~ "Hispanic",
                                       HISPALLP_A == 2 ~ "White only",
                                       HISPALLP_A == 3 ~ "Black/African American only", 
                                       HISPALLP_A == 4 ~ "Asian only",
                                       HISPALLP_A == 5 ~ "AIAN only",
                                       HISPALLP_A == 6 ~ "AIAN and any other group",
                                       HISPALLP_A == 7 ~ "Other single and multiple races",
                                       HISPALLP_A == 97 ~ "Refused",
                                       HISPALLP_A == 98 ~ "Not Ascertained",
                                       HISPALLP_A == 99 ~ "Don't Know")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                      DIBEV_A == 2 ~ "No",
                                      DIBEV_A == 7 ~ "Refused", 
                                      DIBEV_A == 8 ~ "Not Ascertained",
                                      DIBEV_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(EYEEX12M_A =case_when (EYEEX12M_A == 1 ~"Yes",
                                       EYEEX12M_A == 2 ~"No",
                                       EYEEX12M_A == 7 ~"Refused",
                                       EYEEX12M_A == 8 ~"Not Ascertained",
                                       EYEEX12M_A == 9~"Don't Know")) %>%
  filter(DIBEV_A %in% c("Yes"), 
         HISPALLP_A %in% c("Hispanic", "White only", "Black/African American only",
                           "Asian only", "AIAN only", "AIAN and any other group",
                           "Other single and multiple races"),
         EYEEX12M_A%in% c("Yes", "No")) %>%
  dplyr::group_by(EYEEX12M_A) %>%
  dplyr::count(HISPALLP_A) %>%
  dplyr::mutate(perc_eye_exam = (n/sum(n)*100)) %>%
  rename(Respondent_Race = HISPALLP_A,
         Eye_Exam = EYEEX12M_A)

knitr::kable(eye_exam_race_tbl)

eye_exam_race_tbl %>%
  ggplot(aes(x = reorder(Respondent_Race, -n), y = n, fill = Eye_Exam)) + 
  geom_col(position = "fill") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Number of respondents with diabetes that took an eye exam across the races",
       x ="Did the respondent take an eye exam?", 
       y = "Proportion")

eye_exam_race_tbl %>%
  ggplot(aes(x = reorder(Respondent_Race, -n), y = n, fill = Eye_Exam)) + 
  geom_col(position = "dodge2") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Number of respondents with diabetes that took an eye exam across the races",
       x ="Did the respondent take an eye exam?", 
       y = "Count")

```

## 4) Looking at differences in responses based on respondents' county 

### a) Number of respondents that have diabetes across county types 

```{r, message = FALSE, warning = FALSE}

diabetes_county_tbl<- NHIS_Diabetes_Eye %>%
  dplyr::mutate(URBRRL = case_when(URBRRL == 1 ~ "Large central metro",
                                       URBRRL == 2 ~ "Large fringe metro",
                                       URBRRL == 3 ~ "Medium and small metro", 
                                       URBRRL == 4 ~ "Nonmetropolitan")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                      DIBEV_A == 2 ~ "No",
                                      DIBEV_A == 7 ~ "Refused", 
                                      DIBEV_A == 8 ~ "Not Ascertained",
                                      DIBEV_A == 9 ~ "Don't Know")) %>%
  filter(DIBEV_A %in% c("Yes", "No")) %>%
  dplyr::group_by(DIBEV_A) %>%
  dplyr::count(URBRRL) %>%
  dplyr::mutate(perc_diabetes = (n/sum(n)*100)) %>%
  rename(Respondent_County = URBRRL,
         Respondent_Diabetes = DIBEV_A)

kable(diabetes_county_tbl)

diabetes_county_tbl %>%
  ggplot(aes(x = reorder(Respondent_Diabetes, -n), y = n, fill = Respondent_County)) + 
  geom_col(position = "fill") + 
  scale_fill_brewer(palette = "Paired") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of respondents that have diabetes across the counties",
       x ="Does the respondent have diabetes?", 
       y = "Count")

diabetes_county_tbl %>%
  ggplot(aes(x = reorder(Respondent_Diabetes, +n), y = n, fill = Respondent_County)) + 
  geom_col(position = "dodge2") + 
  scale_fill_brewer(palette = "Paired") + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title="Number of respondents that have diabetes across the counties",
       x ="Does the respondent have diabetes?", 
       y = "Count")

```

### b) Number of respondents with diabetes that took an eye exam across county types 

```{r}


eye_exam_race_tbl <- NHIS_Diabetes_Eye %>%
  dplyr::mutate(URBRRL = case_when(URBRRL == 1 ~ "Large central metro",
                                   URBRRL == 2 ~ "Large fringe metro",
                                   URBRRL == 3 ~ "Medium and small metro", 
                                   URBRRL == 4 ~ "Nonmetropolitan")) %>%
  dplyr::mutate(DIBEV_A = case_when(DIBEV_A == 1 ~ "Yes",
                                      DIBEV_A == 2 ~ "No",
                                      DIBEV_A == 7 ~ "Refused", 
                                      DIBEV_A == 8 ~ "Not Ascertained",
                                      DIBEV_A == 9 ~ "Don't Know")) %>%
  dplyr::mutate(EYEEX12M_A =case_when (EYEEX12M_A == 1 ~"Yes",
                                       EYEEX12M_A == 2 ~"No",
                                       EYEEX12M_A == 7 ~"Refused",
                                       EYEEX12M_A == 8 ~"Not Ascertained",
                                       EYEEX12M_A == 9~"Don't Know")) %>%
  filter(DIBEV_A %in% c("Yes"),
         EYEEX12M_A%in% c("Yes", "No")) %>%
  dplyr::group_by(EYEEX12M_A) %>%
  dplyr::count(URBRRL) %>%
  dplyr::mutate(perc_eye_exam = (n/sum(n)*100)) %>%
  rename(Respondent_County = URBRRL,
         Eye_Exam = EYEEX12M_A)

knitr::kable(eye_exam_race_tbl)

eye_exam_race_tbl %>%
  ggplot(aes(x = reorder(Respondent_County, -n), y = n, fill = Eye_Exam)) + 
  geom_col(position = "fill") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Number of respondents with diabetes that took an eye exam across the counties",
       x ="Did the respondent take an eye exam?", 
       y = "Proportion")

eye_exam_race_tbl %>%
  ggplot(aes(x = reorder(Respondent_County, -n), y = n, fill = Eye_Exam)) + 
  geom_col(position = "dodge2") + 
  scale_fill_manual(values = c("coral3", 
                               "cornflowerblue")) +
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_light() +
  theme(text = element_text(size = 10), 
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Number of respondents with diabetes that took an eye exam across the counties",
       x ="Did the respondent take an eye exam?", 
       y = "Count")

```

